Studying public transit API query logs\\ to get an indication of travel flows

### 2016-01-07 paper.tex
Added figure references

%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
%This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Context & % Need
In the field of urban planning, researchers need an indication of how people move between cities. 
Yet, getting statistics of travel flows within public transit systems has proven to be troublesome.
% Task
In order to get an indication of travel flows between cities in Belgium,
we analyzed the query logs of the iRail API, a highly expressive route planning API for the Belgian railways.
We were able to study $\sim$100k to 500k requests for each month between October 2012 and November 2015, which is between $0.56\%$ and $1.66\%$ of the amount of monthly passengers.
% Object & Findings
Using data visualizations, we illustrate the commuting patterns in Belgium and confirm that Brussels, the capital, acts as a central hub. 
The Flemish region appears to be polycentric, while in the Walloon region, everything converges on Brussels.
The findings correspond to the real travel demand, according to experts of the passenger federation Trein Tram Bus.
% Conclusion
We conclude that query logs of route planners are of high importance in getting an indication of travel flows.
% Perspectives
However, better analytics would be acquirable using dedicated HTTP POST requests to capture travel intentions.
%%Redundant according to RV:
%Highly expressive transport data publishing methods such as route planning API exist, as well as data dumps or data fragments.
%In order to be able to gather meaningful logs in all cases, we suggest using a separate POST request containing the entire query.

@@

\begin{figure*}
  \centering
  \includegraphics[width=14cm]{method-c.pdf}\\
  \caption{Each station is color coded to indicate the province it will be aggregated to. Stations in major cities as well as stations outside of Belgium are aggregated to seperate groups.}
  \label{fig:method-b}
\end{figure*}

\begin{figure}[h]
  \centering
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-a.pdf}\\
    \subcaption{} % 4 words max
    %\label{fig:method-a}
  \end{subfigure}%
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-b.pdf}\\
    \subcaption{} % 4words max
    %\label{fig:method-b}
  \end{subfigure}%
  \vspace{5.0pt}% To create a bit of vertical space between the 2 figure captions
  \label{fig:method-a}
  \caption{Spatial aggregation based on provincial administrative boundaries diagrammatically represented. (a) Individual data records are depicted geographically with arrows connecting origins to their respective destinations. (b) The result of aggregation is a set of weighted connections between provinces.}
\end{figure}

Next, the data is aggregated so that province level flow patterns are emphasized.
\todo{Pieter: I can't seem to link the figure here} provides a diagrammatic representation of spatial aggregation on the basis of provincial administrative boundaries.
To compare the volume of flow between rural provincial areas and urbanized zones, train stations in major cities are aggregated seperately. 
If an origin or destination station is a located outside of Belgium, it is considered international travel and aggregated in a distinct group. A diagrammtic breakdown is depicted in \todo{Pieter: Same problem here...} where color and shape is used to demarcate each group.
The following types of flows are observable from the outcome of aggregation:

>>>>

%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
%This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Context & % Need
In the field of urban planning, researchers need an indication of how people move between cities. 
Yet, getting statistics of travel flows within public transit systems has proven to be troublesome.
% Task
In order to get an indication of travel flows between cities in Belgium,
we analyzed the query logs of the iRail API, a highly expressive route planning API for the Belgian railways.
We were able to study $\sim$100k to 500k requests for each month between October 2012 and November 2015, which is between $0.56\%$ and $1.66\%$ of the amount of monthly passengers.
% Object & Findings
Using data visualizations, we illustrate the commuting patterns in Belgium and confirm that Brussels, the capital, acts as a central hub. 
The Flemish region appears to be polycentric, while in the Walloon region, everything converges on Brussels.
The findings correspond to the real travel demand, according to experts of the passenger federation Trein Tram Bus.
% Conclusion
We conclude that query logs of route planners are of high importance in getting an indication of travel flows.
% Perspectives
However, better travel intentions would be acquirable using dedicated HTTP POST requests.
%%Redundant according to RV:
%Highly expressive transport data publishing methods such as route planning API exist, as well as data dumps or data fragments.
%In order to be able to gather meaningful logs in all cases, we suggest using a separate POST request containing the entire query.

@@

\begin{figure*}
  \centering
  \includegraphics[width=14cm]{method-c.pdf}\\
  \caption{Each station is color coded to indicate the province it will be aggregated to. Stations in major cities as well as stations outside of Belgium are aggregated to seperate groups.}
  \label{fig:method-c}
\end{figure*}

\begin{figure}[h]
  \centering
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-a.pdf}\\
    \subcaption{} % 4 words max
    \label{fig:method-a}
  \end{subfigure}%
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-b.pdf}\\
    \subcaption{} % 4words max
    \label{fig:method-b}
  \end{subfigure}%
  \vspace{5.0pt}% To create a bit of vertical space between the 2 figure captions
  \label{fig:method}
  \caption{Spatial aggregation based on provincial administrative boundaries diagrammatically represented. (a) Individual data records are depicted geographically with arrows connecting origins to their respective destinations. (b) The result of aggregation is a set of weighted connections between provinces.}
\end{figure}

Next, the data is aggregated so that province level flow patterns are emphasized.
\cref{fig:method} provides a diagrammatic representation of spatial aggregation on the basis of provincial administrative boundaries.
To compare the volume of flow between rural provincial areas and urbanized zones, train stations in major cities are aggregated seperately. 
If an origin or destination station is a located outside of Belgium, it is considered international travel and aggregated in a distinct group. 
A diagrammatic breakdown is depicted in \cref{fig:method-c} where color and shape is used to demarcate each group.
The following types of flows are observable from the outcome of aggregation:
